---
title: "Report on Care Availability and Mortality Rate Improvements"
output:
  html_document:
    df_print: paged
date: "2025-07-28"
---

```{r}
#| label: Options
#| include: false
#| echo: false
# Set default options for all subsequent code chunks
#knitr::opts_chunk$set(tidy = 'styler') # For demonstration (tidyverse style)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE) # Else
```
```{r}
#| label: setup
#| error: FALSE
#| warning: FALSE
#| message: FALSE
library(tidyverse)
library(here)
```
```{r}
#| label: data
source(here("scripts", "load.R"))
#merged <- readRDS(here("data", "merged.Rds"))
# Minor changes to the data for later use
ind_levels <- c("MNCH_ANC4", "MNCH_SAB")
ind_labels <- c("ANC4 (4+ antenatal care visits)", "SBA (delivery with skilled health personnel)")
pace_levels <- c("Off-track", "On-track")
merged <- merged %>% mutate(
  Indicator = factor(Indicator, levels = ind_levels, labels = ind_labels),
  Pace = factor(Pace, ordered = TRUE)
  )
```

...

## Analysis 

The figure displays the weighted average of maternal care access, comparing countries "on track" to meet their five-year mortality rate goals with those that "need acceleration."
On-track countries show substantially higher rates of care, with access to ANC4 and SBA being approximately 15 and 20 percentage points higher, respectively. This suggests an association between greater access to maternal care and progress in reducing child mortality.
However, this observation should be interpreted with caution. While the better access to cares may improve the mortality rate, it is possible that a confounding factor, such as higher income level and better education, improves both the access to cares and the child mortality rate.
Furthermore, there is significant variance in access rates, particularly within the "need acceleration" group. The wide distribution for this group, with several countries achieving access rates above 80%, indicates that simply having high access does not guarantee a country will be on track.

```{r}
# Calculate the average of the measurement values for each indicator, grouped by the status
mean_both <- merged %>% group_by(Indicator, Pace) %>% 
  summarise(Mean = weighted.mean(Value, w = Births, na.rm = TRUE)) %>% ungroup()
# Plot them as bar graphs (per the requirement, they are shown in a single visualization)
mean_both %>%
  ggplot(aes(x = Pace, y = Mean)) + geom_bar(stat = "identity") +
  #geom_jitter(data = merged, aes(x = Pace, y = Value, size = Births, alpha = 0.5)) +
  facet_wrap(~ Indicator) +
  xlab("Under-five mortality classification") +
  ylab("Average shares of women with the care (%)") +
  labs(title = "Comparison of maternal care access shares (n=152)",
       caption = "Sources: UNICEF Global Data Repository\nData are from 2018-2022")
```

...


## Notes 

I apply for the following positions:

- **Learning and Skills Data Analyst Consultant**– Req. #581598
- **Household Survey Data Analyst Consultant**– Req. #581656
- **Adminisative Data Analyst**– Req. #581696
- **Microdata Harmonization Consultant**– Req. #581699

The report presents strictly a single visualization and a single paragraph, following the instruction.
Other sections like introduction or data discussion are omitted.